---
title: "Global Order Analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```




*Step 1: Import the Global Orders 2016 data set using the read.csv function*

```{r step1, echo=FALSE, message=FALSE, include=FALSE}
# Read CSV into R
globalorder <- read.csv(file="D:/Github/Global Order Analysis/Global Superstore 2016 Orders.csv", header=TRUE, sep = ',')
globalorder$Order.Year<-as.numeric(as.character(globalorder$Order.Year))

```
**The CSV file has been successfully imported**

*Step 2 - Create a horizontal barplot of Total Sales by Market, ordered descending on total sales (i.e. the market with highest*


```{r step2, echo=FALSE, message=FALSE, warning=FALSE}
#create a data frame for total sales of each market
library(dplyr)
order_count <- as.data.frame(globalorder %>% group_by(Market) %>% summarise(Total_Sale=sum(Sales)))
View(order_count)

#Create a horizontal barplot of Total Sales by Market, ordered descending on total sales 
#library(ggplot2)

#convert the data frame to ascending order
order_count<-order_count[order(order_count$Total_Sale,decreasing = FALSE),]

#switch to horizontal bars
barplot(order_count$Total_Sale, 
        names.arg = order_count$Market,
        #xlab = "Total Sales",
        las=1,
        space = 0.75,
        col = c("light grey","light grey","light grey","light grey","light grey","light grey","light grey"),
        horiz = TRUE,
        xlim = range(0,max(order_count$Total_Sale)))
title("Total Sales by Market",adj = 0)
```

*Does any market appear to be an outlier in total sales?*


**The Canada market has extremely low sales as compared to other markets and can be considered as an outlier.**

*Can you visually indicate this market as a separate color from the others?*

```{r step2_outlier, echo=FALSE}
#visually indicate the market which appear to be an outlier
barplot(order_count$Total_Sale, 
        names.arg = order_count$Market,
        #xlab = "Total Sales",
        las=1,
        space = 0.75,
        col = c("dark blue","light grey","light grey","light grey","light grey","light grey","light grey"),
        horiz = TRUE,
        xlim = range(0,max(order_count$Total_Sale)))
title("Total Sales by Market",adj = 0)
#order_count$Order.Year<-as.numeric(as.character(order_count$Order.Year))

```


*Step 3 - Create a line chart total sales by year for each market*

```{r step3, echo=FALSE}
library(dplyr)
year_sale <- as.data.frame(globalorder %>% group_by(Market,Order.Year) %>% summarise(Total_Sale=sum(Sales)),Year=Order.Year)
#year_sale$Order.Year<-as.numeric(as.character(year_sale$Order.Year))
#View(year_sale)

year_sale.Africa<-year_sale[year_sale$Market=="Africa",]
year_sale.APAC<-year_sale[year_sale$Market=="APAC",]
year_sale.Canada<-year_sale[year_sale$Market=="Canada",]
year_sale.EMEA<-year_sale[year_sale$Market=="EMEA",]
year_sale.EU<-year_sale[year_sale$Market=="EU",]
year_sale.LATAM<-year_sale[year_sale$Market=="LATAM",]
year_sale.US<-year_sale[year_sale$Market=="US",]

#find x and y axis range
xrange<-range(year_sale$Order.Year)
yrange<-range(year_sale$Total_Sale)

#plot the range first
plot(xrange, 
     yrange, 
     ylim = range(0,max(year_sale$Total_Sale)),
     las=1,
     xlab="Year",
     ylab="",
     xaxt="n")
title("Total Sales per year for each market",adj = 0)

#modify axes
axis(1, labels = as.character(year_sale$Order.Year), at = as.numeric(year_sale$Order.Year))

#add point markers
points(year_sale.Africa$Order.Year,year_sale.Africa$Total_Sale,type = "p",pch=19)
points(year_sale.APAC$Order.Year,year_sale.APAC$Total_Sale,pch=19,type = "p")
points(year_sale.Canada$Order.Year,year_sale.Canada$Total_Sale,pch=19,type = "p")
points(year_sale.EMEA$Order.Year,year_sale.EMEA$Total_Sale,pch=19, type="p")
points(year_sale.EU$Order.Year,year_sale.EU$Total_Sale,pch=19,type= "p")
points(year_sale.LATAM$Order.Year,year_sale.LATAM$Total_Sale,pch=19,type = "p")
points(year_sale.US$Order.Year,year_sale.US$Total_Sale,pch=19,type = "p")

#add lines
lines(year_sale.Africa$Order.Year,year_sale.Africa$Total_Sale,col = "Red",pch=18,type = "b")
lines(year_sale.APAC$Order.Year,year_sale.APAC$Total_Sale, col = "Dark orange",pch=18,type = "b")
lines(year_sale.Canada$Order.Year,year_sale.Canada$Total_Sale, col= "Dark Blue",pch=18,type = "b")
lines(year_sale.EMEA$Order.Year,year_sale.EMEA$Total_Sale, col="Green",pch=18,type = "b")
lines(year_sale.EU$Order.Year,year_sale.EU$Total_Sale, col = "Hot Pink",pch=18,type="b")
lines(year_sale.LATAM$Order.Year,year_sale.LATAM$Total_Sale, col = "Black",pch=18,type="b")
lines(year_sale.US$Order.Year,year_sale.US$Total_Sale, col="Purple",type = "b", pch=18)

#add text to highlight each market
text(2013,1195000,"APAC",cex = 1.0,adj = -2)
text(2013,1025000,"EU",cex = 1.0,adj = -5)
text(2013,775000,"US",cex = 1.0,adj = -6)
text(2013,635000,"LATAM",cex = 1.0,adj = -2)
text(2013,330000,"EMEA",cex = 1.0,adj = -2.5)
text(2013,245000,"Africa",cex = 1.0,adj = -3)
text(2013,75000,"Canada",cex = 1.0,adj = -2)
```


*Does the same market appear to be an outlier in your line graph as well?*


**Yes, the same market appear to be an outlier in the line graph as well.**

*Can you visually indicate this market as a separate color from the other markets in your graph?*

```{r step3_outlier, echo=FALSE}
library(dplyr)
year_sale <- as.data.frame(globalorder %>% group_by(Market,Order.Year) %>% summarise(Total_Sale=sum(Sales)),Year=Order.Year)
#year_sale$Order.Year<-as.numeric(as.character(year_sale$Order.Year))
#View(year_sale)

#make a data frame for each market and include the year and total sale
year_sale.Africa<-year_sale[year_sale$Market=="Africa",]
year_sale.APAC<-year_sale[year_sale$Market=="APAC",]
year_sale.Canada<-year_sale[year_sale$Market=="Canada",]
year_sale.EMEA<-year_sale[year_sale$Market=="EMEA",]
year_sale.EU<-year_sale[year_sale$Market=="EU",]
year_sale.LATAM<-year_sale[year_sale$Market=="LATAM",]
year_sale.US<-year_sale[year_sale$Market=="US",]

#find x and y axis range
xrange<-range(year_sale$Order.Year)
yrange<-range(year_sale$Total_Sale)

#plot the range first
plot(xrange, 
     yrange, 
     #ylim = range(0,max(year_sale$Total_Sale)),
     las=1,
     xlab="Year",
     ylab="",
     xaxt="n")
title("Market with the lowest Sales in four years",adj = 0)

#modify axes
axis(1, labels = as.character(year_sale$Order.Year), at = as.numeric(year_sale$Order.Year))

#add point markers
points(year_sale.Africa$Order.Year,year_sale.Africa$Total_Sale,type = "p",pch=1)
points(year_sale.APAC$Order.Year,year_sale.APAC$Total_Sale,pch=1,type = "p")
points(year_sale.Canada$Order.Year,year_sale.Canada$Total_Sale,pch=1,type = "p")
points(year_sale.EMEA$Order.Year,year_sale.EMEA$Total_Sale,pch=1, type="p")
points(year_sale.EU$Order.Year,year_sale.EU$Total_Sale,pch=1,type= "p")
points(year_sale.LATAM$Order.Year,year_sale.LATAM$Total_Sale,pch=1,type = "p")
points(year_sale.US$Order.Year,year_sale.US$Total_Sale,pch=1,type = "p")

#add line to indicate the outlier
lines(year_sale.Africa$Order.Year,year_sale.Africa$Total_Sale,col = "light grey",pch=18,type = "b")
lines(year_sale.APAC$Order.Year,year_sale.APAC$Total_Sale, col = "light grey",pch=18,type = "b")
lines(year_sale.Canada$Order.Year,year_sale.Canada$Total_Sale, col= "Dark Blue",pch=18,type = "b",lwd=4)
lines(year_sale.EMEA$Order.Year,year_sale.EMEA$Total_Sale, col="light grey",pch=18,type = "b")
lines(year_sale.EU$Order.Year,year_sale.EU$Total_Sale, col = "light grey",pch=18,type="b")
lines(year_sale.LATAM$Order.Year,year_sale.LATAM$Total_Sale, col = "light grey",pch=18,type="b")
lines(year_sale.US$Order.Year,year_sale.US$Total_Sale, col="light grey",type = "b", pch=18)

#add the text to highlight the market having the lowest sales
text(2013,75000,"Canada",cex = 1.2,adj = -1.5)
```


*Step 4 - Create a box plot of total sales by market*

```{r step4, echo=FALSE}
#create a box plot
#attach databse to R search path so that objects in the database cab be accessed simply using their names
attach(globalorder)
#to compare the sales of different market, separate the sales with the market
boxplot(Sales~Market,
        #ylab="Sales",
        las=1,
        ylim=range(0,1200))
title("Comparison of the Sales of different markets", adj=0)
```

*Does the general pattern you observe match that of the earlier steps 2 and 3?*


**Here, the analysis of the sales have been done on the scale of 0-1200. In this particular scenario the canadian market still has the lowest total sale however, the median sale of Africa, EMEA and US is comparable to that of Canada.**


*What other insights can you draw from your box plot above?*

**For Canadian market, the interquartile range is 145, which means 50% of the sales lie within 25 and 170.** 

*Can you visually indicate this market as a separate color from the others in your boxplot graph?*

```{r step4_outlier, echo=FALSE}
#Highlight the canadian market
boxplot(Sales~Market,
        #ylab="Sales",
        las=1,
        ylim=range(0,1200),
        col= c("Light Grey","Light Grey", "Dark Blue","Light Grey","Light Grey","Light Grey","Light Grey"))
title("Comparison of the Sales of different markets", adj=0)

```

*Step 5 - Determine the relation between the Profit and Sales*

*Not a bar plot, line graph, or box plot*

**The Scatter plot has been used to determine the relation between the Profit and Sales of different products from various Categories. This analysis could be used to further expand the business in Canada which has the lowest Sales.**


```{r step5_outlier, echo=FALSE}
#scatter plot to show relation between Sales and Profit for various categories
#create new data frames for each category having profit and sales
myvars<-names(globalorder) %in% c("Category", "Sales", "Profit")
Tech_sale<-globalorder[which(Category=='Technology'),]
Tech_sale<-Tech_sale[myvars]

off_sale<-globalorder[which(Category=='Office Supplies'),]
off_sale<-off_sale[myvars]

fur_sale<-globalorder[which(Category=='Furniture'),]
fur_sale<-fur_sale[myvars]

#plot the range 
profitrange<-range(min(globalorder$Profit),globalorder$Profit)
salerange<-range(0,globalorder$Sales)
plot(salerange, 
     profitrange, 
     ylim = range(min(globalorder$Profit),max(globalorder$Profit)),
     xlim = range(0,max(globalorder$Sales)),
     las=1,
     xlab="Selling price of a product",
     ylab="Profit earned",
     col="White")
title("Profit Vs Selling Cost for the products of different Categories", adj=0)

#draw different point of each category
points(fur_sale$Sales,fur_sale$Profit,type = "p",pch=17,col="Dark Grey")
points(off_sale$Sales,off_sale$Profit,type = "p",pch=15,col="Dark Golden rod")
points(Tech_sale$Sales,Tech_sale$Profit,type = "p",pch=16,col="Brown")
#draw regression line for technology
regpoint_tech<-lm(Tech_sale$Profit ~ Tech_sale$Sales)
abline(regpoint_tech, col="Brown",lwd=5)

#draw regression line for Furniture
regpoint_fur<-lm(fur_sale$Profit ~ fur_sale$Sales)
abline(regpoint_fur,col="Dark Grey")

#draw regression line for Office Supplies
regpoint_off<-lm(off_sale$Profit ~ off_sale$Sales)
abline(regpoint_off,col="Dark golden rod", lwd=1)

#Add legend
legend("topleft", 
       legend = c("Technology", "Furniture","Office Supplies"), 
       col = c("Brown","Dark Grey","Dark golden rod"), 
       pch = c(16,17,15), 
       bty = "n", 
       pt.cex = 1, 
       cex = 1, 
       text.col = "black",
       title = "Categories")


```


